---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 1"
author: "Rong Guang"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# 1 Preparation

## 1.1 Load packages 

```{r, message = F, warning = F}
library(tidyverse) 
library(corrplot) 
library(psych)
library(GGally) 
```

## 1.2 Load dataset

```{r}
# read the data from the Survo website:
deca <- read.table("https://survo.fi/data/Decathlon.txt", header = TRUE, sep = '\t')
```

## 1.3 Check dataset

### 1.3.1 Check first 15 rows

```{r}
library(knitr)
deca[-1] %>% head(15) %>% t() %>% kable (align = "rrrrrrrrrrrrr") 
```

### 1.3.2 Check dimensions

```{r}
paste(dim(deca) [1], "observations")
paste(dim(deca) [2], "variables")
```

### 1.3.3 Check summary statistics

```{r}
library(finalfit)
kable(ff_glimpse(deca)$Continuous[,c(3:4,6:12)], digits = 2, align = "rrrrrrrrr")
```

## 1.4 Add labels to variables

```{r}
library(expss)
deca <-  apply_labels(deca, Name = "name of athlete",
                            Points = "number of points obtained by the athlete",
                              Run100m = "run 100m in seconds",
                              Longjump = "Long jump in meters",
                              Shotput = "shot-put in meters",
                              Hijump = "High jump in meters",
                              Run400m = "run 400m in seconds",
                              Hurdles = "110 m hurdles in meters",
                              Discus = "discus-throw in meters",
                              Polevlt = "pole-valut in meters",
                              Javelin = "Javelin-throw in meters",
                              Run1500m = "run 1500m in seconds",
                              Height = "height",
                              Weight = "weight")
```

## 1.5 Check labels

```{r}
library(sjlabelled)
get_label(deca)
```


# 2 Data visualization

## 2.1 Scatterplot

### 2.1.1 First day sports

```{r, message = F, warning = F}
#define a function that allows me for more control over ggpairs
#this function produces density plot 
my.fun.density <- function(data, mapping, ...) { #notes are roughly same with above

    ggplot(data = data, mapping = mapping) +
       geom_histogram(aes(y=..density..),
                      color = "black", 
                      fill = "white")+
       geom_density(fill = "#FF6666", alpha = 0.25) +
       theme(panel.grid.major = element_blank(), 
             panel.grid.minor = element_blank(),
             panel.background = element_rect(fill = "#9999CC",
                                             color = "black"))
} 

#define a function that allows me to fine-tune the matrix
my.fun.smooth <- function(data,    #my function needs 3 arguments
                          mapping,
                          method = "loess"){
  ggplot(data = data, #data is passed from ggpairs' arguments
         mapping = mapping)+#aes is passed from ggpairs' arguments
           geom_point(size = 0.3,  #draw points
                      color = "blue")+
           geom_smooth(method = method,  #fit a linear regression
                       size = 0.3, 
                       color = "red")+
           theme(panel.grid.major = element_blank(), #get rid of the grids
                 panel.grid.minor = element_blank(),
                 panel.background = element_rect(fill = "#F0E442", #adjust background
                                                 color = "black"))
} 

#draw it
deca %>% select(Run100m:Run400m) %>% 
  ggpairs(lower = 
          list(continuous = my.fun.smooth), #lower half show points with fitted line
        diag =
          list(continuous = my.fun.density), #diagonal grids show density plots
        title = "Fig1. Relationships among performance of first day sports") + #title
  theme (plot.title = element_text(size = 11,  #adjust title visuals
                                   face = "bold")) 
```

### 2.1.2 Second day sports

```{r, message = F, warning = F}
#draw it
deca %>% select(Hurdles:Run1500m) %>% 
  ggpairs(lower = 
          list(continuous = my.fun.smooth), #lower half show points with fitted line
        diag =
          list(continuous = my.fun.density), #diagonal grids show density plots
        title = "Fig2. Relationships among performance of second day sports") + #title
  theme (plot.title = element_text(size = 11,  #adjust title visuals
                                   face = "bold")) 
```

### 2.1.3 Height and weight

```{r}
deca %>% ggplot(aes(x = Height, y = Weight)) +
  geom_point(shape = 16, size = 5, color = "red", alpha = .4)+
  geom_point(shape = 1, size = 5, color = "black")+
  theme_bw()+
  theme(panel.grid.major = element_blank())+
  labs(title = "Fig3. Scatterplot for height and weight")+
  geom_smooth(method="loess")

```

## 2.2 Visual correlation matrix

### Correlation strength visualized in size

```{r, caption = "Fig4. Visual correlation matrix for the variables"}
deca %>% 
  select(-Name) %>% 
  cor() %>% 
  corrplot(title = "Fig4. Visual correlation matrix for the variables",
           mar=c(0,0,1,0))
```

### Correlation strength shown in correlation coefficients

```{r}
deca %>% select(-Name) %>% cor() %>% 
  corrplot(method = 'number', 
           number.cex = 0.7, 
           number.digits = 1,
           title = "Fig5. Correlation matrix for the variables",
           mar=c(0,0,1,0))
```

# 3 Exploratory factor analysis

## 3.1 Check linearity assumption

Findings were the assumption of multivariate normality was basically met (at least half of the variables were not normal), see Fig 1 and 2. 

## 3.2 Check factoribility

Factorability of the items was examined. Several well-recognized criteria for the factorability of a correlation were used.

### 3.2.1 Correlation matrix

It was observed that all of the 10 items correlated at least 0.3 with at least one other item, suggesting good factorability. See fig 6.

```{r}
deca_perform <- deca %>% 
  select(-c("Name","Points", "Height", "Weight")) 

deca_perform %>% cor() %>% 
  corrplot(method = 'number',
           type = "upper",
           number.cex = 0.7, 
           number.digits = 1,
           title = "Fig6. Correlation matrix for the performance variables",
           mar=c(0,0,1,0))
```

### 3.2.2 KMO test

```{r, fig.width=8}
deca_matrix <- deca_perform %>% 
  cor.plot()
KMO(deca_matrix) 
```

The Kaiser-Meyer-Olkin measure of sampling adequacy was .61, indicating mediocre adequacy according to Kaiser.

### 3.2.3 Bartlett test

```{r}
cortest.bartlett(deca_matrix, nrow(deca_perform))
```

Bartlett’s test of sphericity was significant (p < .05), suggesting that the null hypothesis that our matrix be equal to an identity matrix was objected, and hence good factorability.

## 3.3 Estimate factor structure

The number of factors was estimated. Several well-recognized criteria for exploring factor structure solution were used.

```{r}
fa.parallel(deca_matrix, fa = "fa", fm = "ml", plot = T)
```

The scree plot graphs the eigenvalue against the factor number. From the third factor on, the line is almost flat, meaning the each successive factor is accounting for smaller and smaller amounts of the total variance.

## 3.4 Very simple structure and BIC

```{r}
a <- nfactors(deca_matrix, n=3, rotate = "oblimin", fm = "pa", n.obs=12)
```

The parallel analysis and scree plot based on correlation matrix suggested 3-factor structure, and the VSS was also favourable to a 3-factor interpretation. This was further consolidated by sample size adjusted BIC. While results from the Velicer MAP achieves a minimum of 0.05  with  2  factors, pointing to the optimal number of two factor, and Emprirical BIC suggested a uni-dimension solution. Overall, 3 factor solution was the most universally favored structure, and it was hence adopted in the current task.

## 3.4 Factor analysis

### 3.4.1 Remove items not well explained by factors

```{r}
#factoring the items
deca_fa3 <- fa(deca_perform, nfactors=3, fm="ml", 
             rotate = "varimax", scores = "regression")
#display results
print(deca_fa3$communalities)
```

Variables Polevlt, Javelin, Hijump and Hurdles showed very low commonality (<0.3), suggesting the proportion of their variance cannot be well captured by the factors. They were hence removed.


```{r}
deca_perform_reduced <- deca_perform %>% 
  select(-Polevlt, -Javelin, -Hurdles, -Hijump)
```

```{r}
deca_reduced_fa3 <- fa(deca_perform_reduced, nfactors=3, fm="ml", 
             rotate = "varimax", scores = "regression")

print(deca_reduced_fa3$communalities)
```

Variables Longjump showed very low commonality (<0.3), suggesting the proportion of its variance cannot be well captured by the factors. It was hence removed.

```{r}
deca_perform_reduced <- deca_perform_reduced %>% 
  select(-Longjump)
```

```{r}
deca_reduced_fa3 <- fa(deca_perform_reduced, nfactors=3, fm="pf", 
             rotate = "varimax", scores = "regression")

print(deca_reduced_fa3, digits = 2, sort = TRUE)
```

Then all the items had commonalities >0.3, suggesting they can be well-represented by the factors.

### 3.4.2 Factor structure analysis for the remaining 5 items

(1) scree plot

```{r}
deca_matrix <- deca_perform_reduced %>% 
  cor.plot()
fa.parallel(deca_matrix, fa = "fa", fm = "ml", plot = T)
```

Elbow effect was observed at the 3rd factor.

(2) Very simple structure and BIC

```{r}
deca_reduced_fa3 <- fa(deca_perform_reduced, nfactors=3, fm="ml", 
             rotate = "varimax", scores = "regression")

print(deca_reduced_fa3, digits = 2, sort = TRUE)

nfactors(deca_matrix, n=3, rotate = "oblimin", fm = "pa", n.obs=12)
```

Rotation method was switched to varimax since the performance of these remaining sports can be dependent on different capacity. 

VSS and BCI results suggested a 1 or 2 factor solution. I tested the 2-factor solution.

```{r}
deca_reduced_fa2 <- fa(deca_perform_reduced, nfactors=2, fm="ml", 
             rotate = "varimax", scores = "regression")

print(deca_reduced_fa2$loadings,cutoff = 0.3, sort=TRUE)

fa.diagram(deca_reduced_fa2, cut = 0, main = "Factor Analysis, orthogonal rotation")
```

Shot-put, Discus and Run1500m predominantly loaded together. Run100m and Run400m loaded together. Though Run400m also negatively loaded together with the other factor, the variability explained is quite smaller than when it loads with Run100m 

As such, I defined the factors in the following way. F1-Shotput, Discus and Run1500m-captures a latent factor of strength; F2-Run100m and Run1500m-captures a latent factor of explosiveness. 

# 4 Generate factor scores

```{r}
scores <- as.data.frame(deca_reduced_fa2$scores) %>%
  mutate(Name = deca$Name) %>%
  rename(fa_strength = ML1, 
         fa_explosiveness = ML2)
#join into the dataset
deca_new <- left_join(deca, scores)
```


# 5 Regression

## 5.1 Check distribution

```{r}
factor.labs <- c("Explosiveness","Strength")
names(factor.labs) <- c("fa_explosiveness", "fa_strength")


scores %>% 
  select(contains("fa_")) %>%  
  pivot_longer(everything(), 
               names_to = "factor", 
               values_to = "score") %>% 
  ggplot(aes(x = score))+
  geom_histogram(bins=7, color = "black", fill = "lightblue")+
  facet_wrap(~factor, labeller = labeller(factor = factor.labs))+
  theme(panel.background = element_rect(fill = "white",
                                        color = "black"),
        strip.background = element_rect(color = "black",
                                        fill = "coral"),
        strip.text = element_text(color= "white",
                                  size = 10,
                                  face = "bold"))
```

## 5.2 Check scatterplot

```{r}
library(ggrepel)
deca_new %>% 
  ggplot(aes(y = fa_strength, x = fa_explosiveness, label = Name)) +
  geom_point() + geom_text_repel()
```

```{r, message = F, warning = F}
#draw it
deca_new %>% select(Points, contains("fa_")) %>% 
  ggpairs(lower = 
          list(continuous = my.fun.smooth), #lower half show points with fitted line
        diag =
          list(continuous = my.fun.density)) + #diagonal grids show density plots) 
  theme (plot.title = element_text(size = 11,  #adjust title visuals
                                   face = "bold")) 
```

## 5.3 Predicting points

### 5.3.1 model

```{r}
lm_deca <- lm(Points ~ fa_strength + fa_explosiveness, data = deca_new)
summary(lm_deca)
```

### 5.3.2 finetue model

```{r}
lm_deca <- lm(Points ~ fa_explosiveness, data = deca_new)
summary(lm_deca)
```

# 5.3.3 Model Diagnostic

```{r}
library(ggfortify)
autoplot(lm_deca)
```

# 6 Confirmatory factor analysis (Bonus)

# 6.1 Sepcify the model

```{r}
install.packages("lavaan")
library(lavaan)

deca.cfa <- ' Strength =~ Discus + Shotput
              Explosiveness =~ Run100m + Run400m
              Endurance =~ Run1500m '
```


# 6.2 Fit the model and display the output

```{r}
fit <- cfa(deca.cfa, data = deca)
summary(fit, fit.measures = T)
```

# 6.3 

```{r}


deca.cfa <- ' Strength =~ Discus + Shotput + Run1500m + Polevlt
              Explosiveness =~ Run100m + Run400m + Hijump
              Skill =~ Longjump + Hurdles '
```


```{r}
fit <- cfa(deca.cfa, data = deca)
summary(fit, fit.measures = T)
```


```{r}
deca.cfa <- ' Strength =~ Discus + Shotput + Polevlt
              Explosiveness =~ Run100m + Run400m + Hijump + Hurdles +Javelin
              Endurance =~ Run1500m'
```


```{r}
fit <- cfa(deca.cfa, data = deca)
summary(fit, fit.measures = T)
```

```{r}
```

```{r}
deca.cfa <- ' Strength =~ Discus + Shotput + Run1500m
              Explosiveness =~ Run100m + Run400m'
```


```{r}
fit <- cfa(deca.cfa, data = deca)
summary(fit, fit.measures = T)
```


```{r}
```


```{r}
```


```{r}
```


