---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 2"
author: "Rong Guang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Exercise 2.1

Specify and test the hypothesis given on the page 1 of the lecture material.

Draw conclusions based on the $\chi^2$ statistic and the CFI, TLI, RMSEA, and SRMR indices.

What can you say about the parameter estimates?

Visualize the model.

## 1.1 Read in the data set

Start by downloading the data file from Moodle to Project folder.

```{r}
library(tidyverse)#data wrangling 
library(readr)# read data into r
orig_data <- read_csv("ASC7INDM.CSV", show_col_types = FALSE) 
```

## 1.2 Subset the data set

Subset the variables for analysis and name it as sc (Self-concept).

```{r}
# Select the variables for use
sc <- orig_data %>% dplyr::select(starts_with("SDQ2N")) # namming logic: sc = self-concept
```

## 1.3 Inspect the data

Have a quick overview of the data.

```{r}
glimpse(sc)
```

The data set includes 16 variables from 265 observations. All the variables are numeric.

```{r}
library(purrr)
lapply(sc, function(x)unique(x)) %>% nest()
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```










