---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 2"
author: "Rong Guang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, warning = F, message=F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message=F)
```

# 1 Exercise 2.1

Specify and test the hypothesis given on the page 1 of the lecture material.

Draw conclusions based on the $\chi^2$ statistic and the CFI, TLI, RMSEA, and SRMR indices.

What can you say about the parameter estimates?

Visualize the model.

## 1.1 Read in the data set

Start by downloading the data file from Moodle to Project folder.

```{r, warning = F, message=F}
library(tidyverse)#data wrangling 
library(readr)# read data into r
orig_data <- read_csv("ASC7INDM.CSV", show_col_types = FALSE) 
```

## 1.2 Write functions

```{r}
unique.levels <-  function(sc){
  values <- lapply(sc, function(x)sort(unique(x))) 
for(x in 1:ncol(sc)){
  a <- paste(c("Variable ", 
               names(values)[x], 
               " has values of ", 
               paste(values[[x]], 
                     collapse = ",")), 
             collapse = "")
  print(a)
  }
}
```


## 1.3 Subset the data set

Subset the variables for analysis and name it as sc (Self-concept).

```{r}
# Select the variables for use
sc <- orig_data %>% dplyr::select(starts_with("SDQ2N")) # namming logic: sc = self-concept
```

## 1.4 Inspect the data

Have a quick overview of the data.

```{r}
glimpse(sc)
```

The data set includes 16 variables from 265 observations. All the variables are numeric. Next, I examined the unique values of each variables.

```{r}
unique.levels(sc)
```

For each variable, the values distribute from 1 to 6. 

# 2 Explore the data

## 2.1 Descriptive statistics

```{r, warning = F, message=F}
library(kableExtra)#publication-ready table
library(psych)#for function "describe"
sc.ds <- sc %>%  #sc.ds = self-concept descriptive statistics
  describe(IQR = T) %>%
  select(mean, median, sd, range, se, IQR)

#print the descriptive statistics table
sc.ds %>% 
  kable(booktabs=T,
        longtable=T,
        digits = 2,
        caption = "Descriptive dtatistics of selected variables",
        linesep = "") %>% 
  add_header_above(c("", "centralized tendency" = 2, "dispersion tendency" = 4)) %>% 
  kable_styling(latex_options = c("striped","repeat_header")) %>% 
  column_spec(1, width = "3cm", bold = T, color = "red")
#sc.ds %>%
#  tab_df(digits = 2, 
#        alternate.rows = T,
#        title = "Table 1. Descriptive dtatistics of selected variables",
#        CSS = list(css.centralign='text-align: right;'))
```

## 2.2 Visualization

### 2.2.1 Histogram

```{r, fig.width = 6, fig.height=5, warning = F, message=F}
sc %>% 
  pivot_longer(everything()) %>%  #longer format
  ggplot(aes(x = value)) + #x axis used variable "value" (a default of pivot)
  geom_histogram(binwidth = 1, aes(y = ..density..), #match ys of density and histogram plots
                 color = "black",  fill = "#9999CC")+  # adjust aesthetics for hist
  geom_density(fill = "pink", alpha = 0.25)+ #adjust aesthetics for density plot
  facet_wrap(~name, scales = "free") + #wrap by name variable
  theme(panel.grid.major = element_blank(), #get rid of the  grids
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white",#adjust the background
                                        color = "black"),
        strip.background = element_rect(color = "black",#adjust the strips aes
                                        fill = "steelblue"),
        strip.text = element_text(size =8, color = "white"), #adjust strip text
        axis.title.x = element_text(size = 3), #adjust the x text
        axis.title.y = element_text(size = 3), # adjust the y text
        plot.title = element_text(size = 12, face = "bold"))+ #adjust the title
  labs(title = "Figure 1 Distribution of selected items") #title it
```

### 2.2.2 Correlation plot

```{r}
library(GGally)
ggcorr(sc, 
       geom = "blank", 
       label = TRUE, 
       hjust = 0.85, 
       color = "red", 
       face = "bold", 
       method = c("pairwise","spearman"),
       digits = 2,
       label_size = 2.5,
       label_round = 2) +
  geom_point(size = 9, 
             aes(color = "red", 
                 alpha = abs(coefficient) > 0.3)) +
  scale_alpha_manual(values = c("TRUE" = 0.3, "FALSE" = 0)) +
    geom_point(size = 10, 
               aes(color = "green", alpha = abs(coefficient) > 0.6)) +
  scale_alpha_manual(values = c("TRUE" = 0.5, "FALSE" = 0)) +
  guides(color = FALSE, 
         alpha = FALSE) +
  labs(title = "Figure 2. Spearman correlation matrix of the selected items",
       caption = 
         "Red circles indicates correlation coefficient ≥ 0.5; gree circle indicates ≥ 0.3")
```
It is found that each variable correlated with at least one of the other variable with a spearman correlation coefficient ≥0.3, except for item SDQ2N46 and SDQ2N34. 

```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```










