---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 2"
author: "write your name here"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CFA & self-concept (SC)

## Exercise 2.1

Specify and test the hypothesis given on the page 1 of the lecture material.

Draw conclusions based on the $\chi^2$ statistic and the CFI, TLI, RMSEA, and SRMR indices.

What can you say about the parameter estimates?

Visualize the model.

### Read in the data set:

Start by downloading the data file from Moodle to your Project folder!

```{r}
library(tidyverse)
library(readr)

orig_data <- read_csv("ASC7INDM.CSV", show_col_types = FALSE) 

# we will only use a subset of the data here:
SCdata <- orig_data %>% dplyr::select(starts_with("SDQ2N"))
glimpse(SCdata)
```

### Explore the data (always do that!):

```{r}
library(psych)
library(dplyr)
library(knitr)
library(tidyr)
library(corrplot)

# these are just examples - you can well do other things, too!

# basic statistics:
SCdata %>% describe() %>% 
  as.data.frame() %>%
  select(mean, sd, min, max) %>%
  kable(digits = 2)

# histograms:
SCdata %>% pivot_longer(cols = everything()) %>%
  ggplot(aes(x = value)) + 
  geom_histogram(binwidth = 1) +
  facet_wrap(~name, nrow = 4, ncol = 4)

# correlation matrix:
SCdata %>% cor() %>% round(digits = 2)

# correlation plot:
SCdata %>% cor() %>% corrplot()

```

### Define and estimate a CFA model:

```{r}
library(lavaan) # install.packages("lavaan")

# Define a CFA model using the lavaan package:
# NOTE: with the model definitions in lavaan syntax you have to
# SELECT all the code up to ' and then press Ctrl+Enter / Cmd+Enter
# when activating operations individually.

model1 <- '# CFA model of self-concept (SC):
           GSC =~ SDQ2N01 + SDQ2N13 + SDQ2N25 + SDQ2N37
           ASC =~ SDQ2N04 + SDQ2N16 + SDQ2N28 + SDQ2N40
           ESC =~ SDQ2N10 + SDQ2N22 + SDQ2N34 + SDQ2N46
           MSC =~ SDQ2N07 + SDQ2N19 + SDQ2N31 + SDQ2N43
          '

# Estimate the model using the data defined earlier:
cfa1 <- cfa(model1, data = SCdata)

# Numerical summary of the model:
summary(cfa1, fit.measures = TRUE, standardized = TRUE)

```

### Visualize the CFA model:

```{r}
library(semPlot) # install.packages("semPlot")

# Path model based on the model object (TRY AND MODIFY!)
semPaths(cfa1) # default options

# LISREL style (introduced by K. Jöreskog in the 1970s - still the standard):
semPaths(cfa1, style = "lisrel")

# "some" more options added:
semPaths(cfa1, style = "lisrel", layout = "tree2", what = "path", whatLabels = "name",
    intercepts = FALSE, residuals = TRUE, thresholds = FALSE, reorder = FALSE,
    rotation = 2, 
    latents = c("MSC","ESC", "ASC" ,"GSC"),
    sizeLat = 10, sizeLat2 = 5,
    manifests = rev(colnames(SCdata)),
    sizeMan = 10, sizeMan2 = 2
    )

```

## Exercise 2.2

Specify and test these two additional hypotheses (again draw conclusions based on the $\chi^2$ statistic and the CFI, TLI, RMSEA, and SRMR indices):

- *Hypothesis 2:* SC is a two-factor structure consisting of GSC and ASC (so that the four GSC measures load onto the GSC and all other onto the ASC).

- *Hypothesis 3:* SC is a one-factor structure.

Visualize the models and compare them with the four-factor model analyzed in Exercise 2.1.

```{r}

# (copy and modify the R codes given above!)

```

