---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 3"
author: "write your name here"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CFA & teacher burnout

## Exercise 3.1

Specify and test the hypothesis given on the pages 2 and 3 of the lecture material.

Use 1) ML estimator, 2) MLM estimator.

Compare the fit indices and draw conclusions concerning the model fit.

Visualize the model.

### Read in the data set:

Start by downloading the data file from Moodle to your Project folder!

```{r}
library(tidyverse)
library(readr)

MBIdata <- read_csv("ELEMM1.CSV", show_col_types = FALSE) 
glimpse(MBIdata)
```

### Explore the data (always do that!):

```{r}

# (copy and modify the R codes given in Assignment 2)

```

### Define and estimate a CFA model:

Note: you must first complete the model definition, it is not ready yet!

```{r}
library(lavaan)

# Define a CFA model using the lavaan (Latent Variable Analysis) syntax:
# see https://lavaan.ugent.be/tutorial/syntax1.html

model1 <- '
# CFA model for the burnout:
# EE: EmotionalExhaustion
# DP: Depersonalization
# PA: PersonalAccomplishment
  EE =~ ITEM1 + ...
  DP =~ ITEM5 + ...
  PA =~ ITEM4 + ...
'

# Estimate the model with the default (ML) estimator:
cfa1 <- cfa(model1, data = MBIdata)

# Numerical summary of the model:
summary(cfa1, fit.measures = TRUE, standardized = TRUE)

# Use a robust (MLM) estimator:
cfa2 <- cfa(model1, data = MBIdata, estimator = "MLM")

# Numerical summary of the model:
summary(cfa2, fit.measures = TRUE, standardized = TRUE)
```

Example of creating a table for comparing two models (here, ML and MLM):

```{r}
fit_ml <- fitMeasures(cfa1, c("chisq", "cfi", "tli", "rmsea", "rmsea.ci.lower", "rmsea.ci.upper"), output = "matrix")
fit_mlm <- fitMeasures(cfa2, c("chisq.scaled", "cfi.robust", "tli.robust", "rmsea.robust", "rmsea.ci.lower.robust", "rmsea.ci.upper.robust"), output = "matrix")
ML <- as.data.frame(t(fit_ml))
MLM <- as.data.frame(t(fit_mlm))
colnames(ML) <- colnames(MLM) <- c("Chisq", "CFI", "TLI", "RMSEA", "CI lower", "CI upper")
fit_table <- rbind(ML, MLM)
rownames(fit_table) <- c("ML", "MLM")

kable(fit_table, digits = 3, align = "lcccc", booktabs = TRUE,
             caption = "Comparison of fit indices between ML and MLM estimators")
```

### Visualize the CFA model:

```{r}
library(semPlot)

# (copy and modify the R codes given in Assignment 2)

```

## Exercise 3.2

Continue with post hoc model fitting (exploratory approach).

Proceed **step by step** following the guidelines given in the lecture material, i.e., implement the modifications **one at a time**, testing and studying each step. See (and report) how the fit improves and which parameters are suggested to be modified. Please be careful! There will (always) be a lot of suggestions...

You may also test what happens to your model, if you go 'too far' (that is, further than in the lecture material and the book)... *but please come back!* ;-)

Draw the graph of the final model and present its fit indices and the essential parameter estimates (both unstandardized and standardized).

```{r}

# (copy and modify the R codes given earlier)

# You will get the list of the MIs (Modification Indices) as follows (here, cfa2 is a fit object created by the cfa() function)
modindices(cfa2, standardized = TRUE, sort. = TRUE)

# another way, by saving the result to the object MI and listing only the 10 largest ones:
MI <- modindices(cfa2, standardized = TRUE, sort. = TRUE)
head(MI, n = 10)

# still another way, listing only the 10 largest and displaying them as a table:
modindices(cfa2, standardized = TRUE, sort = TRUE, maximum.number = 10) %>%
  kable(digits = 3, align = "rclccccc")

```

