---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 4"
author: "Rong Guang"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, cache = T, fig.align = 'center')
```

**\textcolor{blue}{The texts that reflect my understanding have been highlighted in} \textcolor{red}{red color}.** 

# Task description

The first section is task description, which is copied from the assignment5.rmd. It is for communicating with future "me". Please skip it.

## Exercise 5.1

Specify and estimate the initial baseline models for the two groups. 

Present a brief summary of the model fit and make the first step of the modification by including **(exceptionally, at the same time!)** all the four parameters known to be required for improving the model fit of both models.

Fine-tune the models step by step following the guidelines given in the lecture material, i.e., implement the modifications **(as usually, one change at a time)** testing and studying each step. 

Present the final baseline models of each group and draw the graphs

# Preparation

##Read in the data set:

Start by downloading the **two data files** from Moodle to your Project folder!

```{r}
#install the necessary pakages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(here, 
               expss, 
               tidyverse, 
               janitor,
               knitr, 
               qualtRics, 
               arules, 
               arulesViz, 
               sjlabelled,
               DT,
               stringr,
               labelled,
               ggstatsplot,
               ggcorplot)

library(tidyverse)
library(readr)

#This week's file name
latest.name1 <- "MBIELM1.CSV"
latest.name2 <- "MBISEC1.CSV"
#read in the data
mbi.elm <-  #elementary school
  read_csv(
    file.path(
      here(),
      'data',
      latest.name1
      )
    )

mbi.sec <- #secondary school
  read_csv(
    file.path(
      here(),
      'data',
      latest.name2
      )
    )

```

## Write functions

To control length of reports, codes already shown in the previous homework were not showing in the current report. Yet they are available in .rmd report.

### to generate CFA results with improved readability

```{r, echo = F}
#goodness of fit indicators for ml
cfa.summary.mlm.a <- function(fit){
  options(scipen = 999)
  cfa.measure <- fitMeasures(fit,    #obtain specified measured.
                            c("chisq.scaled", 
                              "df.scaled", 
                              "pvalue.scaled", 
                              "cfi.scaled", 
                              "tli.scaled",
                              "rmsea.scaled",
                              "rmsea.pvalue.scaled",
                              "srmr_bentler",
                              "chisq.scaling.factor")) 
  names(cfa.measure) <- c("chi square", "df", "p value", "CFI", "TLI", "RMSEA", "RMSEA p value", "SRMR", "CSF")
  #turn named vector to data frame
cfa.tab.a <- cfa.measure %>%  
  tibble(name= names(cfa.measure), value = cfa.measure) %>% # vector to df
  select(Measure = name, Value = value) %>%  #select and rename columns
  mutate(Value = round(as.numeric(Value),3)) 
}
#factor loading
cfa.summary.b <- function(fit, fa.num, item.num, estimator){
  options(scipen = 999)
  #factor loading
  cfa.tab.b <- parameterEstimates(fit, standardized=TRUE) %>% # obtain estimates
  filter(op == "=~") %>%  #select "is measured by" rows
  select('Latent Factor'=lhs, #left hand side column
         Indicator=rhs, #right hand side column
         B=est, #estimates
         SE=se, #standard error
         Z=z, #z statistics
         'p-value'=pvalue, #p value
         Beta=std.all) %>%  
  kable(digits = 3, #rounded to 3
        format="markdown", #Latex markdown
        booktabs=TRUE, #Latex booktabs
        caption=paste("Factor Loadings for",fa.num,"factor CFA model estimated by ", estimator)) %>% #caption
  kable_styling(latex_options = "striped") %>% #gray every other row
  row_spec(0, background = "#9999CC") # color the first row
  cfa.tab.b
  }

#Variance
cfa.summary.c <- function(fit, fa.num, item.num, estimator){
  options(scipen = 999)
  #Variance
  type <- rep(c("Residual variance", "Total variance"), 
            time = c(item.num, fa.num)) #create a new row clarifying types of variance

variance <- parameterEstimates(fit, standardized=TRUE) %>% #obtain estimates
  filter(op == "~~") #select "is correlated with" rows
variance <- variance[1:sum(item.num,fa.num),] #subset 1:18 rows (variance row)
variance <- cbind(type, variance) #add column
cfa.tab.c <- variance %>%select(Type = type, #select and rename variables
                   Indicator=rhs, #right hand side column
                   B=est, #estimates
                   SE=se,#standard error
                   Z=z, #z statistics
                   'p-value'=pvalue, #p value
                   Beta=std.all) %>% 
  kable(digits = 3, #rounded
        format="markdown",  #Latex markdown
        booktabs=TRUE, #Latex booktabs
        caption=paste("Variances for", fa.num, "factor model estimated by ", estimator)) %>% #caption
  kable_styling(latex_options = "striped") %>% # gray every other row
  row_spec(0, background = "#9999CC") # color the variable row
  cfa.tab.c
}

#Covariance
cfa.summary.d <- function(fit, fa.num, item.num, estimator){
  options(scipen = 999)
  #covariance
  variance <- parameterEstimates(fit, standardized=TRUE) %>%
  filter(op == "~~")
  covar.num = (fa.num+(fa.num-1))/2
variance <- variance[sum(item.num,fa.num,1):sum(item.num,fa.num,covar.num),]
type <- paste(variance$lhs, "with", variance$rhs) 
variance <- cbind(type, variance)
rownames(variance) <- NULL
cfa.tab.d <- variance %>%select(Type=type, 
                   B=est, 
                   SE=se,
                   Z=z, 
                   'p-value'=pvalue, 
                   Beta=std.all) %>% 
  kable(digits = 3, 
        format="markdown", 
        booktabs=TRUE, 
        caption=paste("Covariances for", fa.num, 
                      "factor model estimated by ", 
                      estimator)) %>% 
  kable_styling(latex_options = "striped") %>% 
  row_spec(0, background = "#9999CC")
  cfa.tab.d
}

```

### to generate a function for correlation matrix with numbers

```{r, echo = F}
mymatrix <- function(data, fig.num = 3){
  library(GGally)
ggcorr(data, 
       geom = "blank", 
       label = TRUE, 
       hjust = 0.9, 
       color = "red", 
       face = "bold", 
       method = c("pairwise","pearson"),
       digits = 2,
       size= 2.5,
       label_size = 2.5,
       label_round = 2,
       layout.exp =1) +
  geom_point(size = 7, 
             aes(color = "steelblue", 
                 alpha = abs(coefficient) > 0.3)) +
  scale_alpha_manual(values = c("TRUE" = 0.4, 
                                "FALSE" = 0)) +
    geom_point(size = 8, 
               aes(color = "red", 
                   alpha = abs(coefficient) > 0.6)) +
  scale_alpha_manual(values = c("TRUE" = 0.4, 
                                "FALSE" = 0)) +
  guides(color = FALSE, 
         alpha = FALSE) +
  labs(title = paste("Figure ", fig.num," Pearson correlation matrix of the selected items"),
       caption = 
         " Red circles indicates the absolute of correlation coefficient >= 0.6 
        green circle indicates >= 0.3")+
  theme(plot.title = element_text(size = 12,
                                  face = "bold",
                                  hjust = 0.5),
        plot.caption = element_text(color = "red"))
}

```

### to generate a function for histogram overlapping with density plot

```{r, echo = F}
corr.density <- function(data, fig.num = 1){
  data %>% 
  pivot_longer(everything()) %>%  #longer format
  ggplot(aes(x = value)) + #x axis used variable "value" (a default of pivot)
  geom_histogram(binwidth = 1, aes(y = ..density..), #match ys of density and histogram plots
                 color = "black",  fill = "#9999CC")+  # adjust aesthetics for hist
  geom_density(fill = "pink", alpha = 0.25)+ #adjust aesthetics for density plot
  facet_wrap(~name, scales = "free", ncol =4) + #wrap by name variable
  theme(panel.grid.major = element_blank(), #get rid of the  grids
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white",#adjust the background
                                        color = "black"),
        strip.background = element_rect(color = "black",#adjust the strips aes
                                        fill = "steelblue"),
        strip.text = element_text(size =8, color = "white"), #adjust strip text
        axis.title.x = element_text(size = 3), #adjust the x text
        axis.title.y = element_text(size = 3), # adjust the y text
        plot.title = element_text(size = 12, 
                                  face = "bold",
                                  hjust = 0.5))+ #adjust the title
  labs(title = paste("Figure ", fig.num," Distribution of selected items")) #title it
  }

```

### to generate a function for violin overlapping with box plot

```{r, echo = F}
violin.box <- function(data, fig.num = 2){
  mbi.long <- data %>% pivot_longer(everything(), names_to = "item", values_to = "score")

mbi.long %>% 
  ggplot(aes(x = item, y = score)) +
  geom_violin(trim=F, fill = "#9999CC") +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust =1),
        axis.title = element_text(size = 12),
        panel.background = element_rect(fill = "white", color = "black"),
        plot.title = element_text(face="bold",
                                  hjust = 0.5),
        axis.title.x = element_blank())+
  labs(x = "Item", 
       y = "Score",
       title = paste("Figure 2 ", fig.num, " Violin plot of the selected items"))+
  geom_boxplot(width = 0.1, fill = "white")
}
```

### To generate a function describing continuous data set

```{r, echo=F}
descriptive <- function(data){
  library(finalfit)
library(kableExtra)
inspect.table <- ff_glimpse(data)$Continuous
inspect.table$label <- NULL
inspect.table %>% 
  mutate('Q1Q3' = paste(quartile_25, 
                        quartile_75, 
                        sep = " ~ ")) %>% 
  select(n, 
         'n of NA' = missing_n, 
         'Mean' = mean, 
         'Median' = median,
         'SD' = sd, 
         'Min' = min, 
         'Max' = max,
         'Q1~Q3' = Q1Q3) %>%
  kable(booktabs = T,  
        align = "r",
        longtable = T,
        linesep = "",
        caption = "Descriptive statistics for measurements") %>% 
  add_header_above(c(" ", 
                     " " = 2,
                     "Central tendency" = 2, 
                     "Dispersion tendency" = 4)) %>% 
  kable_styling(latex_options = c("striped", 
                                  "repeat_header")) %>% 
  column_spec(1, width = "3cm")
}
```

### Write a function describing continuous data set

```{r, echo=F}
descriptive <- function(data){
  library(finalfit)
library(kableExtra)
inspect.table <- ff_glimpse(data)$Continuous
inspect.table$label <- NULL
inspect.table %>%
  mutate('Q1Q3' = paste(quartile_25,
                        quartile_75,
                        sep = " ~ ")) %>%
  select(n,
         'n of NA' = missing_n,
         'Mean' = mean,
         'Median' = median,
         'SD' = sd,
         'Min' = min,
         'Max' = max,
         'Q1~Q3' = Q1Q3) %>%
  kable(booktabs = T,
        align = "r",
        longtable = T,
        linesep = "",
        caption = "Descriptive statistics for measurements") %>%
  add_header_above(c(" ",
                     " " = 2,
                     "Central tendency" = 2,
                     "Dispersion tendency" = 4)) %>%
  kable_styling(latex_options = c("striped",
                                  "repeat_header")) %>%
  column_spec(1, width = "3cm")
}
```

### Write a function for histogram overlapping with density plot

```{r, echo = F}
corr.density <- function(data, fig.num = 1, group){
  data %>%
  pivot_longer(everything()) %>%  #longer format
  ggplot(aes(x = value)) + #x axis used variable "value" (a default of pivot)
  geom_histogram(binwidth = 1, aes(y = ..density..), #match ys of density and histogram plots
                 color = "black",  fill = "#9999CC")+  # adjust aesthetics for hist
  geom_density(fill = "pink", alpha = 0.25)+ #adjust aesthetics for density plot
  facet_wrap(~name, scales = "free", ncol =5) + #wrap by name variable
  theme(panel.grid.major = element_blank(), #get rid of the  grids
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white",#adjust the background
                                        color = "black"),
        strip.background = element_rect(color = "black",#adjust the strips aes
                                        fill = "steelblue"),
        strip.text = element_text(size =8, color = "white"), #adjust strip text
        axis.title.x = element_text(size = 3), #adjust the x text
        axis.title.y = element_text(size = 3), # adjust the y text
        plot.title = element_text(size = 12,
                                  face = "bold",
                                  hjust = 0.5))+ #adjust the title
  labs(title = paste("Figure", fig.num," Distribution of selected items for", group)) #title it
  }
```

### Write a function to generate dot distribution plot

```{r}
dot.dist <- 
  function(data, type, title){
    data |>
      t() |> 
      as.data.frame() %>% 
      mutate(Item = rownames(.)) |> 
      rowwise() |> 
      mutate(Median = eval(parse(text = type))(V1:V580)) |> 
      ggstatsplot::ggdotplotstats(
        point.args = list(color = "red", size = 3, shape = 13),
        xlab = paste(type, "ratings"),
        title = title,
        x = Median,
        y = Item
      )
    }
```

### Write a fuction to generate correlation matrix with statistical test

```{r}
mycor <- 
  function(data, cols, title){
  mbi.elm |> 
      select(all_of(cols)) |> 
      ggstatsplot::ggcorrmat(
        colors = c("#B2182B", "white", "#4D4D4D"),
        title = "(a) Items on emotional exhaustion, 
        elementary school teacher",
        matrix.type  = "lower"
      )
    }
```

# Inspect the data

## Distribution

```{r,fig.width= 8, fig.height = 13}
#generate the plots, by subgroup of teachers
p.dist.elm <- 
  corr.density(
    mbi.elm, 
    fig.num = "1(a)", 
    group = "elementary school teacher"
    )
p.dist.sec <- 
  corr.density(
    mbi.sec, 
    fig.num = "1(b)",
    group = "secondary school teacher"
    )
#print the plot
library(patchwork); p.dist.elm/p.dist.sec
```

 
```{r,fig.width= 8, fig.height = 5}
#generate plot by subgroups of teachers
p.dot.elm <- 
  dot.dist(
    data = mbi.elm, 
    type = "median", 
    title = "(a) Elementary school teacher"
    )
p.dot.sec <- 
  dot.dist(
    data = mbi.sec, 
    type = "median", 
    title = "(b) Secondary school teacher"
    )
#plot layout
patchwork <- p.dot.elm|p.dot.sec
#print the plot with a genral title
patchwork+plot_annotation(
    title = 
      'Figure 2 Distributions of median rating for each item',
    theme = 
      theme(plot.title = 
              element_text(
                size = 16,
                face = "bold",
                vjust = -1.5,
                hjust =0.5)
            )
    )
```


```{r,fig.width= 10, fig.height = 15}
fa.ee <- c("ITEM1", "ITEM3", "ITEM6", "ITEM8", "ITEM13", "ITEM14", "ITEM16", "ITEM20")
fa.dp <- c("ITEM5", "ITEM10", "ITEM11", "ITEM15", "ITEM22")
fa.pa <- c("ITEM4", "ITEM7", "ITEM9", "ITEM12", "ITEM17", "ITEM18", "ITEM19", "ITEM21")
#generate 6 plots, 3 factors X 2 subgroups of teachers
p.cor.elm.ee <- 
       mycor(
         data= mbi.elm, 
         cols = fa.ee, 
         "(a) Items on emotional exhaustion, 
         elementary school teacher"
         )
p.cor.sec.ee <- 
       mycor(
         data = mbi.sec, 
         cols = fa.ee, 
         "(b) Items on emotional exhaustion, 
          secondary school teacher"
         )
p.cor.elm.dp <- 
       mycor(
         data = mbi.elm, 
         cols = fa.dp, 
         "(c) Items on depersonalization,
          elementary school teacher"
         )
p.cor.sec.dp <- 
       mycor(
         data = mbi.sec, 
         cols = fa.dp, 
         "(d) Items on depersonalization,
          secondary school teacher"
         )
p.cor.elm.pa <- 
       mycor(
         data = mbi.elm, 
         cols = fa.pa, 
         "(e) Items on personal accomplishment,
         secondary school teacher"
         )
p.cor.sec.pa <- 
       mycor(
         data = mbi.sec ,
         cols = fa.pa, 
         "(f) Items on personal accomplishment,
          secondary school teacher"
         )
#plot sub-figure layout
patchwork <- 
  p.cor.elm.ee/p.cor.elm.dp/p.cor.elm.pa|p.cor.sec.ee/p.cor.sec.dp/p.cor.sec.pa 
#print the plot with a gernal title
patchwork+
  plot_annotation(
    title = 
      'Figure 3 Correlalogram for items on each factor for two groups of teachers',
    theme = 
      theme(plot.title = 
              element_text(
                size = 16,
                face = "bold",
                vjust = -1.5,
                hjust =0.5)
            )
    )

```

# Testing the factorial invariance of MBI inventory between elementary and secondary school teachers  

## Define and estimate initial models for both subgroups

The model under test in this initial multi-group application is the same postulated three-factor structure of the MBI that was tested in the previous assignments. 

### Define the baseline model 

```{r}
library(lavaan)
# Define a CFA model using the lavaan (Latent Variable Analysis) syntax:
# see https://lavaan.ugent.be/tutorial/syntax1.html
BLmodel <- '
# CFA model for the burnout, the baseline model:
    EE =~ ITEM1 + ITEM2 + ITEM3 + ITEM6 + ITEM8 + 
          ITEM13 + ITEM14 + ITEM16 + ITEM20
    DP =~ ITEM5 + ITEM10 + ITEM11 + ITEM15 +ITEM22
    PA =~ ITEM4 + ITEM7 + ITEM9 + ITEM12 + 
          ITEM17 + ITEM18 + ITEM19 + ITEM21
          '
```

It is important to note that measuring instruments are often group specific in the way they operate, and, thus, it is possible that baseline models may not be completely identical across groups

### Estimate factorial validity 

(1) Estimate factorial validity for the elementary teacher subgroup

```{r}
cfa.elm <- 
  cfa(
    BLmodel, 
    data = mbi.elm,  
    estimator = "MLM",
    mimic = "Mplus"
    )
```

(2) Estimate factorial validity for the secondary teacher subgroup

```{r}
cfa.sec <- 
  cfa(
    BLmodel, 
    data = mbi.sec,  
    estimator = "MLM",
    mimic = "Mplus"
    )
```

### Evaluate model fit 

```{r}
library(knitr);library(kableExtra)
BL.elm.fit <- cfa.summary.mlm.a(cfa.elm) |> t() |> as.data.frame()
BL.sec.fit <- cfa.summary.mlm.a(cfa.sec) |> t() |> as.data.frame()
BL.both <- rbind(BL.elm.fit[2,], BL.sec.fit[2,]) 
names(BL.both) <- BL.elm.fit[1,]
rownames(BL.both) <- c("Elementary School Teachers",
                       "Secondary School Tearchers")

BL.both <- BL.both |> 
  rename(p = 'p value',
         p2 = 'RMSEA p value',
         chi = 'chi square') |> 
  mutate(df = as.numeric(df) |> round(0),
         p = case_when(
           as.numeric(p) < 0.001 ~ "<0.001",
           as.numeric(p) >= 0.001 ~ p
           ),
         p2 = case_when(
           as.numeric(p2) < 0.001 ~ "<0.001",
           as.numeric(p2) >= 0.001 ~ p2
           )
         ) |>
  mutate('Chi square (df, p)' = 
           paste0(chi, "(", df,", ", p, ")"),
         'RMSEA(p)'           = 
           paste0(RMSEA, "(", p2, ")"
                  )
         ) |> 
  select('Chi square (df, p)', 
         CFI, TLI, 
         'RMSEA(p)', 
         SRMR, 
         'CSF*'= CSF
         ) 

BL.both |> 
  kable(booktabs = T, 
        #format = "markdown", 
        caption = 
          "Fit indices for two subgroups, basline models",
        align = "r"
        ) |> 
  kable_styling(full_width = T) |> 
  footnote(symbol = 
             "Chi square scaling factor"
           ) |>
  column_spec(1, width = "3.5cm") |> 
  column_spec(2, width = "4cm")|> 
  column_spec(3, width = "1cm")|> 
  column_spec(4, width = "1cm")|> 
  column_spec(5, width = "2.5cm")|> 
  column_spec(6, width = "1cm") |> 
  column_spec(7, width = "1cm") 
```

Goodness-of-fit statistics for this baseline model (three factor) reveals that the indices are less than optimal for both elementary (MLM χ2[206] = 826.573; CFI = 0.857; RMSEA = 0.072 ; SRMR = 0.068) and secondary (MLM χ2[206] = 999.359; CFI = 0.836; RMSEA = 0.075; SRMR = 0.077) levels


```{r}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```




```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

