---
title: "COS-D419 Factor Analysis and Structural Equation Models 2023, Assignment 4"
author: "Rong Guang"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, cache = T, fig.align = 'center')
```

**\textcolor{blue}{The texts that reflect my understanding have been highlighted in} \textcolor{red}{red color}.** 

# Task description

The first section is task description, which is copied from the assignment5.rmd. It is for communicating with future "me". Please skip it.

## Exercise 5.1

Specify and estimate the initial baseline models for the two groups. 

Present a brief summary of the model fit and make the first step of the modification by including **(exceptionally, at the same time!)** all the four parameters known to be required for improving the model fit of both models.

Fine-tune the models step by step following the guidelines given in the lecture material, i.e., implement the modifications **(as usually, one change at a time)** testing and studying each step. 

Present the final baseline models of each group and draw the graphs

# Preparation

##Read in the data set:

Start by downloading the **two data files** from Moodle to your Project folder!

```{r}
#install the necessary pakages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(here, 
               expss, 
               tidyverse, 
               janitor,
               knitr, 
               qualtRics, 
               arules, 
               arulesViz, 
               sjlabelled,
               DT,
               stringr,
               labelled,
               ggstatsplot,
               ggcorplot)

library(tidyverse)
library(readr)

#This week's file name
latest.name1 <- "MBIELM1.CSV"
latest.name2 <- "MBISEC1.CSV"
#read in the data
mbi.elm <-  #elementary school
  read_csv(
    file.path(
      here(),
      'data',
      latest.name1
      )
    )

mbi.sec <- #secondary school
  read_csv(
    file.path(
      here(),
      'data',
      latest.name2
      )
    )

```

## Write functions

To control length of reports, codes already shown in the previous homework were not showing in the current report. Yet they are available in .rmd report.

### Write a function describing continuous data set

```{r, echo=F}
descriptive <- function(data){
  library(finalfit)
library(kableExtra)
inspect.table <- ff_glimpse(data)$Continuous
inspect.table$label <- NULL
inspect.table %>%
  mutate('Q1Q3' = paste(quartile_25,
                        quartile_75,
                        sep = " ~ ")) %>%
  select(n,
         'n of NA' = missing_n,
         'Mean' = mean,
         'Median' = median,
         'SD' = sd,
         'Min' = min,
         'Max' = max,
         'Q1~Q3' = Q1Q3) %>%
  kable(booktabs = T,
        align = "r",
        longtable = T,
        linesep = "",
        caption = "Descriptive statistics for measurements") %>%
  add_header_above(c(" ",
                     " " = 2,
                     "Central tendency" = 2,
                     "Dispersion tendency" = 4)) %>%
  kable_styling(latex_options = c("striped",
                                  "repeat_header")) %>%
  column_spec(1, width = "3cm")
}
```

### Write a function for histogram overlapping with density plot

```{r, echo = F}
corr.density <- function(data, fig.num = 1, group){
  data %>%
  pivot_longer(everything()) %>%  #longer format
  ggplot(aes(x = value)) + #x axis used variable "value" (a default of pivot)
  geom_histogram(binwidth = 1, aes(y = ..density..), #match ys of density and histogram plots
                 color = "black",  fill = "#9999CC")+  # adjust aesthetics for hist
  geom_density(fill = "pink", alpha = 0.25)+ #adjust aesthetics for density plot
  facet_wrap(~name, scales = "free", ncol =5) + #wrap by name variable
  theme(panel.grid.major = element_blank(), #get rid of the  grids
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white",#adjust the background
                                        color = "black"),
        strip.background = element_rect(color = "black",#adjust the strips aes
                                        fill = "steelblue"),
        strip.text = element_text(size =8, color = "white"), #adjust strip text
        axis.title.x = element_text(size = 3), #adjust the x text
        axis.title.y = element_text(size = 3), # adjust the y text
        plot.title = element_text(size = 12,
                                  face = "bold",
                                  hjust = 0.5))+ #adjust the title
  labs(title = paste("Figure", fig.num," Distribution of selected items for", group)) #title it
  }
```

### Write a function to generate dot distribution plot

```{r}
dot.dist <- 
  function(data, type, title){
    data |>
      t() |> 
      as.data.frame() %>% 
      mutate(Item = rownames(.)) |> 
      rowwise() |> 
      mutate(Median = eval(parse(text = type))(V1:V580)) |> 
      ggstatsplot::ggdotplotstats(
        point.args = list(color = "red", size = 3, shape = 13),
        xlab = paste(type, "ratings"),
        title = title,
        x = Median,
        y = Item
      )
    }
```

### Write a fuction to generate correlation matrix with statistical test

```{r}
mycor <- 
  function(data, cols, title){
  mbi.elm |> 
      select(all_of(cols)) |> 
      ggstatsplot::ggcorrmat(
        colors = c("#B2182B", "white", "#4D4D4D"),
        title = "(a) Items on emotional exhaustion, 
        elementary school teacher",
        matrix.type  = "lower"
      )
    }
```

# Inspect the data

## Distribution

```{r,fig.width= 8, fig.height = 13}
#generate the plots, by subgroup of teachers
p.dist.elm <- 
  corr.density(
    mbi.elm, 
    fig.num = "1(a)", 
    group = "elementary school teacher"
    )
p.dist.sec <- 
  corr.density(
    mbi.sec, 
    fig.num = "1(b)",
    group = "secondary school teacher"
    )
#print the plot
library(patchwork); p.dist.elm/p.dist.sec
```

 
```{r,fig.width= 8, fig.height = 5}
#generate plot by subgroups of teachers
p.dot.elm <- 
  dot.dist(
    data = mbi.elm, 
    type = "median", 
    title = "(a) Elementary school teacher"
    )
p.dot.sec <- 
  dot.dist(
    data = mbi.sec, 
    type = "median", 
    title = "(b) Secondary school teacher"
    )
#plot layout
patchwork <- p.dot.elm|p.dot.sec
#print the plot with a genral title
patchwork+plot_annotation(
    title = 
      'Figure 2 Distributions of median rating for each item',
    theme = 
      theme(plot.title = 
              element_text(
                size = 16,
                face = "bold",
                vjust = -1.5,
                hjust =0.5)
            )
    )
```


```{r,fig.width= 10, fig.height = 15}
fa.ee <- c("ITEM1", "ITEM3", "ITEM6", "ITEM8", "ITEM13", "ITEM14", "ITEM16", "ITEM20")
fa.dp <- c("ITEM5", "ITEM10", "ITEM11", "ITEM15", "ITEM22")
fa.pa <- c("ITEM4", "ITEM7", "ITEM9", "ITEM12", "ITEM17", "ITEM18", "ITEM19", "ITEM21")
#generate 6 plots, 3 factors X 2 subgroups of teachers
p.cor.elm.ee <- 
       mycor(
         data= mbi.elm, 
         cols = fa.ee, 
         "(a) Items on emotional exhaustion, 
         elementary school teacher"
         )
p.cor.sec.ee <- 
       mycor(
         data = mbi.sec, 
         cols = fa.ee, 
         "(b) Items on emotional exhaustion, 
          secondary school teacher"
         )
p.cor.elm.dp <- 
       mycor(
         data = mbi.elm, 
         cols = fa.dp, 
         "(c) Items on depersonalization,
          elementary school teacher"
         )
p.cor.sec.dp <- 
       mycor(
         data = mbi.sec, 
         cols = fa.dp, 
         "(d) Items on depersonalization,
          secondary school teacher"
         )
p.cor.elm.pa <- 
       mycor(
         data = mbi.elm, 
         cols = fa.pa, 
         "(e) Items on personal accomplishment,
         secondary school teacher"
         )
p.cor.sec.pa <- 
       mycor(
         data = mbi.sec ,
         cols = fa.pa, 
         "(f) Items on personal accomplishment,
          secondary school teacher"
         )
#plot sub-figure layout
patchwork <- 
  p.cor.elm.ee/p.cor.elm.dp/p.cor.elm.pa|p.cor.sec.ee/p.cor.sec.dp/p.cor.sec.pa 
#print the plot with a gernal title
patchwork+
  plot_annotation(
    title = 
      'Figure 3 Correlalogram for items on each factor for two groups of teachers',
    theme = 
      theme(plot.title = 
              element_text(
                size = 16,
                face = "bold",
                vjust = -1.5,
                hjust =0.5)
            )
    )

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```




```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

